---
title: "Quality checks"
output: html_document
author: "Vicente Arrona"
date: "2024-12-13"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

These are the anonymised datasets by country and the complete dataset

```{r import datasets}
ds <- readRDS("data/data_complete.rds")
ds_sp <- readRDS("data/data_Spain.rds")
ds_cr <- readRDS("data/data_Czech_Republic.rds")
ds_nl <- readRDS("data/data_Netherlands.rds")
ds_it <- readRDS("data/data_Italy.rds")


## Codebook for reference
cb <- 
  openxlsx::read.xlsx("ext/all_codebook.xlsx", sheet = 1)
```

Each contributor could use this script or create his/her own to run quality
checks. Please make sure to use both code and text to provide as much
information as possible an ensure replicability. Please also use the package
todor to flag things to do (`todor::todo()`) or to fix (`todor::fixme()`).

```{r req packs}
library(tidyverse)
library(ggpubr) # arranges ggplots
library(visdat)
library(naniar)
library(knitr)
library(gtsummary)
library(skimr)
library(kableExtra) #Saves tables made with kable
library(patchwork) 
library(scales)
library(haven)
```

We noticed that mp25d (First wave) had different response codifications for 0-3
and 0-4, one is purely quantitative ("No, no problems", "Yes, one problem"...)
and the other is quantitative in a temporal manner ("Not at all", "Less than
once a month"...), that makes this NOT comparable (Check "heroes_codebook.xlsx"
to validate this information).

```{r remove mp25d}
ds <- 
  ds |> 
  select(-mp25d) # Take out mp25d while keeping the rest.
```

On the other hand, mp25d_2and3 (2nd and 3rd waves) has 1 extra level for Spain
data, what can be seen in the summary table is that level "0" is at 0%, while
for the other countries (Except Italy) level "5" is at 0%. Checking the scripts,
we realized that when recoding, Spain data took out the "0" value, and defined
the levels 1-5, as for the others, these are defined 0-4. This means that the
spanish dataset had a displacement of the data, resulting on being the only one
with responses at level "5" and non at level "0", needing recoding.
Unfortunately, Italy has no data for this variable, while it does have it for
the previously deleted mp25d, making it difficult for analysis.

```{r recoding mp25d_2and3 for spain}
table(ds$mp25d_2and3, ds$country_of_residence) # To check this issue.

ds <- 
  ds |>
  mutate(mp25d_2and3 = as.numeric(as.character(mp25d_2and3))) |> # To number
  mutate(mp25d_2and3 = case_when(
    mp25d_2and3 == 1 & country_of_residence == "ES" ~ 0,    # Move levels
    mp25d_2and3 == 2 & country_of_residence == "ES" ~ 1,
    mp25d_2and3 == 3 & country_of_residence == "ES" ~ 2,
    mp25d_2and3 == 4 & country_of_residence == "ES" ~ 3,
    mp25d_2and3 == 5 & country_of_residence == "ES" ~ 4,
    TRUE ~ mp25d_2and3)) |> 
  mutate(mp25d_2and3 = factor(mp25d_2and3)) # To factor again

# To be sure, check the values:
levels(ds$mp25d_2and3)

table(ds$mp25d_2and3, ds$country_of_residence) # Should maintain no. of obs.
```

The variable ra56 has 6 levels, with only Spain having data on the 6th one,
"heroes_codebook" describes level "6" as "Not applicable" (It is not defined in
"all_codebook"), we have to recode those answers to be NA values, as the other
countries (Except IT, who doesn't have responses of this) already recoded this
way.

```{r recode level 6 ra56 as NA for Spain}
table(ds$ra56, ds$country_of_residence)

ds <- 
  ds |>
    mutate(ra56 = as.numeric(as.character(ra56))) |> # To number value. 
    mutate(ra56 = case_when(
    ra56 == 6 & country_of_residence == "ES" ~ NA,
    TRUE ~ ra56)) |> 
  mutate(ra56 = factor(ra56)) # To factor again

table(ds$ra56, ds$country_of_residence)
```

We want to look at missing data trends, by wave and country

```{r missing data by wave}
ds_w1 <- 
  ds |> 
  filter(wave == 1)

ds_w2 <- 
  ds |> 
  filter(wave == 2)

ds_w3 <- 
  ds |> 
  filter(wave == 3)

# Czech Republic

vis_miss(ds_w1 |> 
           filter(country_of_residence == "CZ") |> 
           select(1:103)) +
  ggplot2::labs(
    title = "Missing data (Wave 1): Czech R. (Individual variables)")

vis_miss(ds_w1 |> 
           filter(country_of_residence == "CZ") |> 
           select(104:182)) +
  ggplot2::labs(
    title = "Missing data (Wave 1): Czech R. (Ecological variables)")

vis_miss(ds_w2 |> 
           filter(country_of_residence == "CZ") |> 
           select(1:103)) +
  ggplot2::labs(
    title = "Missing data (Wave 2): Czech R. (Individual variables)")

vis_miss(ds_w2 |> 
           filter(country_of_residence == "CZ") |> 
           select(104:182)) +
  ggplot2::labs(
    title = "Missing data (Wave 2): Czech R. (Ecological variables)")

vis_miss(ds_w3 |> 
           filter(country_of_residence == "CZ") |> 
           select(1:103)) +
  ggplot2::labs(
    title = "Missing data (Wave 3): Czech R. (Individual variables)")

vis_miss(ds_w3 |> 
           filter(country_of_residence == "CZ") |> 
           select(104:182)) +
  ggplot2::labs(
    title = "Missing data (Wave 3): Czech R. (Ecological variables)")


# Italy

vis_miss(ds_w1 |> 
           filter(country_of_residence == "IT") |> 
           select(1:103)) +
  ggplot2::labs(
    title = "Missing data (Wave 1): Italy (Individual variables)")
         

vis_miss(ds_w1 |> 
           filter(country_of_residence == "IT") |> 
           select(104:182)) +
  ggplot2::labs(
    title = "Missing data (Wave 1): Italy (Ecological variables)")
 
# Italy does not have data for 2nd and 3rd waves.

# Netherlands

vis_miss(ds_w1 |> 
           filter(country_of_residence == "NL") |> 
           select(1:103)) +
  ggplot2::labs(
    title = "Missing data (Wave 1): Netherlands (Individual variables)")

vis_miss(ds_w1 |> 
           filter(country_of_residence == "NL") |> 
           select(104:182)) +
  ggplot2::labs(
    title = "Missing data (Wave 1): Netherlands (Ecological variables)")

vis_miss(ds_w2 |> 
           filter(country_of_residence == "NL") |> 
           select(1:103)) +
  ggplot2::labs(
    title = "Missing data (Wave 2): Netherlands (Individual variables)")

vis_miss(ds_w2 |> 
           filter(country_of_residence == "NL") |> 
           select(104:182)) +
  ggplot2::labs(
    title = "Missing data (Wave 2): Netherlands (Ecological variables)")

vis_miss(ds_w3 |> filter(country_of_residence == "NL") |> select(1:103)) +
  ggplot2::labs(title = "Missing data (Wave 3): Netherlands (Individual variables)")


vis_miss(ds_w3 |> filter(country_of_residence == "NL") |> select(104:182)) +
  ggplot2::labs(title = "Missing data (Wave 3): Netherlands (Ecological variables)")

# Spain

vis_miss(ds_w1 |> 
           filter(country_of_residence == "ES") |> 
           select(1:103)) +
  ggplot2::labs(
    title = "Missing data (Wave 1): Spain (Individual variables)")
 

vis_miss(ds_w1 |> 
           filter(country_of_residence == "ES") |> 
           select(104:182)) +
  ggplot2::labs(
    title = "Missing data (Wave 1): Spain (Ecological variables)")

vis_miss(ds_w2 |> 
           filter(country_of_residence == "ES") |> 
           select(1:103)) +
  ggplot2::labs(
    title = "Missing data (Wave 2): Spain (Individual variables)")


vis_miss(ds_w2 |> 
           filter(country_of_residence == "ES") |> 
           select(104:182)) +
  ggplot2::labs(
    title = "Missing data (Wave 2): Spain (Ecological variables)")


vis_miss(ds_w3 |> 
           filter(country_of_residence == "ES") |> 
           select(1:103)) +
  ggplot2::labs(
    title = "Missing data (Wave 3): Spain (Individual variables)")

vis_miss(ds_w3 |> 
           filter(country_of_residence == "ES") |> 
           select(104:182)) +
  ggplot2::labs(
    title = "Missing data (Wave 3): Spain (Ecological variables)")

# Per wave, all countries.

# Wave 1
ds_w1 |> 
    miss_var_summary() |> 
    print(n = 172)

# Wave 2

ds_w2 |> 
    filter(country_of_residence != "IT") |>  # Take out Italy
    miss_var_summary() |> 
    print(n = 172)

# Wave 3

ds_w3 |> 
    filter(country_of_residence != "IT") |>  # Take out Italy
    miss_var_summary() |> 
    print(n = 172)
```

Those variables with 100% missing data need to be deleted.

```{r delete var with 100% missing}
na_percentage <- colMeans(is.na(ds))
na_threshold <- 1
ds <- ds[, na_percentage <= na_threshold]
dim(ds)
```

Seeing the proportion of missing data and number of responses per wave and
country.

```{r missing and n count}
ds <- 
  ds |> 
  mutate(na_count = rowSums(is.na(ds)),
         na_perc = 
           rowSums(is.na(ds)) / 
           (rowSums(is.na(ds)) + rowSums(!is.na(ds)))*100,
         na_quartile = case_when(  
           na_perc < 25 ~ "q1",
           na_perc >= 25 & na_perc < 50 ~ "q2",
           na_perc >= 50 & na_perc < 75 ~ "q3",
           na_perc >= 75 ~ "q4"),
         na_q_country = case_when(
           na_quartile == "q1" & 
             country_of_residence == "CZ" ~ "q1_na_CZ",
           na_quartile == "q2" & 
             country_of_residence == "CZ" ~ "q2_na_CZ",
           na_quartile == "q1" & 
             country_of_residence == "IT" ~ "q1_na_IT",
           na_quartile == "q2" & 
             country_of_residence == "IT" ~ "q2_na_IT",
           na_quartile == "q1" & 
             country_of_residence == "NL" ~ "q1_na_NL",
           na_quartile == "q2" & 
             country_of_residence == "NL" ~ "q2_na_NL",
           na_quartile == "q1" & 
             country_of_residence == "ES" ~ "q1_na_ES",
           na_quartile == "q2" & 
             country_of_residence == "ES" ~ "q2_na_ES",
         )
       )

ds |> 
  select(wave,
         na_quartile,
         country_of_residence,
         na_q_country) |> 
  tbl_summary(by = wave)
```

Finally, generate a file for further use.

```{r generate file}
write_csv(ds, "data/data_complete_clean.csv")

write_rds(ds, "data/data_complete_clean.rds")

write_sav(ds, "data/data_complete_clean.sav")
```
